\documentclass[spanish]{textolivre}

% metadata
\journalname{Texto Livre}
\thevolume{17}
%\thenumber{1} % old template
\theyear{2024}
\receiveddate{\DTMdisplaydate{2024}{3}{20}{-1}}
\accepteddate{\DTMdisplaydate{2024}{5}{4}{-1}}
\publisheddate{\today}
\corrauthor{Gema Alcaraz Mármol}
\articledoi{10.1590/1983-3652.2024.51693}
%\articleid{NNNN} % if the article ID is not the last 5 numbers of its DOI, provide it using \articleid{} commmand 
% list of available sesscions in the journal: articles, dossier, reports, essays, reviews, interviews, editorial
\articlesessionname{articles}
\runningauthor{Alcaraz Mármol}
%\editorname{Leonardo Araújo} % old template
\sectioneditorname{Hugo Heredia Ponce}
\layouteditorname{João Mesquita}

\title{Aprendizaje de lengua extranjera a través de herramientas digitales de corpus. Actitudes y autoeficacia en estudiantes universitarios del grado de Educación Primaria}
\othertitle{Aprendizagem de línguas estrangeiras através de ferramentas de corpus digital. Atitudes e autoeficácia em estudantes universitários do ensino básico}
\othertitle{Foreign language learning through digital corpus tools. Attitudes and self-efficacy in university students of Primary Education teacher training}

\author[1]{Gema Alcaraz Mármol~\orcid{0000-0001-7703-3829}\thanks{Email: \href{mailto:gema.alcaraz@uclm.es}{gema.alcaraz@uclm.es}}}
\affil[1]{Universidad de Castilla-La Mancha,  Departamento de Filología Moderna, Toledo, España.}

\addbibresource{article.bib}
\usepackage{array}

\begin{document}
\maketitle
\begin{polyabstract}
\begin{abstract}
El presente estudio tiene como objetivo explorar la
autoeficacia y actitudes sobre el uso de dos herramientas digitales de
corpus por parte de un grupo de estudiantes de inglés como lengua
extranjera que cursa el Grado en Educación Primaria en la Universidad de
Castilla-La Mancha. Los participantes realizaron una serie de
actividades en el contexto de la metodología basada en datos utilizando
como recurso dos herramientas de corpus: SKELL y BNC. Tras la
realización de las actividades y la manipulación de las herramientas de
corpus, los participantes respondieron un cuestionario donde debían
expresar su grado de acuerdo con cada una de las afirmaciones que
componían dicho cuestionario. El cuestionario hace referencia a aspectos
sobre la dificultad de uso, el nivel de utilidad, o la aplicación fuera
del aula de las herramientas digitales de corpus; también se pregunta a
los participantes por las actividades de aprendizaje basado en datos
para la adquisición de distintos aspectos de la lengua inglesa. Los
resultados apuntan a una actitud positiva y un nivel considerable de
autoeficacia hacia el corpus. Finalmente, se anima a adoptar este tipo
de metodología además de a una formación específica del profesorado para
su correcta y óptima aplicación.

\keywords{Autoeficacia \sep Aprendizaje basado en datos \sep Corpus \sep
	Actitudes \sep Lengua extranjera}
\end{abstract}

\begin{portuguese}
\begin{abstract}
O presente estudo tem como objetivo explorar a auto-eficácia e
as atitudes face à utilização de duas ferramentas de \textit{corpus} digital por parte de um grupo de estudantes de inglês como língua estrangeira que
frequentam uma licenciatura em Educação Primária na Universidade de
Castilla-La Mancha. Os participantes realizaram uma série de atividades
no contexto da metodologia orientada para os dados, utilizando duas
ferramentas de corpus: SKELL e BNC. Após a realização das atividades e
a manipulação das ferramentas de \textit{corpus}, os participantes responderam a um questionário no qual tinham de expressar o seu grau de concordância com cada uma das afirmações que compunham o questionário. O questionário refere-se a aspectos de dificuldade de utilização, nível de utilidade ou aplicação fora da sala de aula das ferramentas de \textit{corpus} digital; os participantes são também questionados sobre atividades de aprendizagem baseadas em dados para a aquisição de diferentes aspectos da língua inglesa. Os resultados apontam para uma atitude positiva e um nível considerável de auto-eficácia em relação ao corpus. Por último,
incentiva-se a adoção deste tipo de metodologia, bem como a formação
específica de professores para a sua aplicação correta e optimizada.

\keywords{Auto-eficácia \sep Aprendizagem baseada em dados \sep \textit{Corpus} \sep Atitudes \sep Língua estrangeira}
\end{abstract}
\end{portuguese}

\begin{english}
\begin{abstract}
The present study aims to explore the self-efficacy and
attitudes towards the use of two digital corpus tools by a group of
students of English as a foreign language studying for a degree in
Primary Education at the University of Castilla-La Mancha. The
participants carried out a series of activities in the context of the
data-driven methodology using two corpus tools: SKELL and BNC. After
carrying out the activities and manipulating the corpus tools, the
participants answered a questionnaire in which they had to express their
degree of agreement with each of the statements that made up the
questionnaire. The questionnaire refers to aspects such as difficulty of
use, level of usefulness, or application outside the classroom of the
digital corpus tools; participants are also asked about data-driven
learning activities for the acquisition of different aspects of the
English language. The results point to a positive attitude and a
considerable level of self-efficacy towards the corpus. Finally, the
adoption of this type of methodology is encouraged as well as specific
teacher training for its correct and optimal application.


\keywords{Self-efficacy \sep Data-driven learning \sep Corpus \sep Attitudes \sep Foreign language}
\end{abstract}
\end{english}
\end{polyabstract}

\input{sec-introduccion.tex}
\input{sec-marcoteorico.tex}
\input{sec-objetivodelainvestigacion}
\input{sec-metodologia.tex}
\input{sec-resultadosydiscusion.tex}
\input{sec-conclusion.tex}



\printbibliography\label{sec-bib}
%conceptualization,datacuration,formalanalysis,funding,investigation,methodology,projadm,resources,software,supervision,validation,visualization,writing,review

\end{document}
