\documentclass[english]{textolivre}

% metadata
\journalname{Texto Livre}
\thevolume{17}
%\thenumber{1} % old template
\theyear{2024}
\receiveddate{\DTMdisplaydate{2024}{1}{8}{-1}}
\accepteddate{\DTMdisplaydate{2024}{5}{2}{-1}}
\publisheddate{\DTMdisplaydate{2024}{8}{5}{-1}}
\corrauthor{Willian Fernandes Araujo}
\articledoi{10.1590/1983-3652.2024.49440}
%\articleid{NNNN} % if the article ID is not the last 5 numbers of its DOI, provide it using \articleid{} commmand 
% list of available sesscions in the journal: articles, dossier, reports, essays, reviews, interviews, editorial
\articlesessionname{articles}
\runningauthor{Araujo and Pires}
%\editorname{Leonardo Araújo} % old template
\sectioneditorname{Daniervelin Pereira}
\layouteditorname{João Mesquita}

\title{Algorithmic literacy: a bibliographic review and debate on education}
\othertitle{Educação para os algoritmos: levantamento bibliográfico e debate sobre o conceito de literacia algorítmica}

\author[1]{Willian Fernandes Araujo~\orcid{0000-0002-3271-6690}\thanks{Email: \href{mailto:willianfaraujo@gmail.com}{willianfaraujo@gmail.com}}}

\author[2]{Fernanda Pires de Sá~\orcid{0000-0001-6172-7594}\thanks{Email: \href{mailto:fernanda.pires@uab.cat}{fernanda.pires@uab.cat}}}

\affil[1]{Universidade de Santa Cruz do Sul, Programa de Pós-Graduação em Educação, Departamento de Gestão de Negócios e Comunicação, Santa Cruz do Sul, RS, Brazil.}

\affil[2]{Universitat Autònoma de Barcelona, Departamento de Comunicación Audiovisual y Publicidad, Bellaterra, Barcelona, Spain.}

\addbibresource{article.bib}

\begin{document}
\maketitle
\begin{polyabstract}
\begin{abstract}
This study aimed to map the field of research on algorithmic literacy (AL) in Portuguese, Spanish, and English. The growing importance and presence of algorithms in various facets of social life motivated this research. To achieve this, four axes of investigation were used: 1) definition of conceptual proposals, 2) identification of emerging pedagogical techniques and strategies in these debates, 3) description of the functions and domains of these different forms of knowledge, and 4) reflection on the challenges and difficulties revealed in the relevant literature. The methodology used was a systematic literature review in the Scopus and Web of Science databases, using the descriptors 'Literacy AND Algorithmic', 'algorithmic OR algorithm AND literacy OR literacies'. Subsequently, 21 publications were analyzed following a systematic selection process. Initially, the analysis indicates a significant increase in work in recent years (2021 and 2022), the predominance of publications in English, and the diversity of publications across different domains. Critical empowerment, awareness, and understanding of algorithms emerge as key themes in the debate, but there is considerable variation in the pedagogical approaches adopted to encourage critical engagement with algorithmic systems.

\keywords{Datafication \sep Platformization \sep Algorithmic systems \sep Education}
\end{abstract}

\begin{portuguese}
\begin{abstract}
O estudo tem como objetivo mapear o campo de investigação sobre literacia algorítmica nas línguas portuguesa, espanhola e inglesa. A crescente importância e presença dos algoritmos em diversas facetas da vida social motivou esta pesquisa. Para isso, foram utilizados quatro eixos de investigação: 1) definição de propostas conceptuais, 2) identificação de técnicas e estratégias pedagógicas emergentes nestes debates, 3) descrição das funções e domínios destas diferentes formas de conhecimento e 4) reflexão sobre os desafios e dificuldades revelados na literatura. A metodologia utilizada foi uma revisão sistemática da literatura nas bases de dados Scopus e Web of Science, utilizando os descritores 'Literacy AND Algorithmic', 'algorithmic OR algorithm AND literacy OR literacies'. Posteriormente, 21 publicações foram analisadas após um fluxo sistemático de seleção. Inicialmente, a análise indica um aumento significativo de trabalhos nos últimos anos (2021 e 2022), a predominância de publicações em língua inglesa e a diversidade de publicações em diferentes domínios. A capacitação crítica, a consciencialização e a compreensão dos algoritmos surgem como temas-chave no debate, mas existe uma variação considerável nas abordagens pedagógicas adotadas para encorajar o envolvimento crítico com os sistemas algorítmicos.

\keywords{Dataficação \sep Plataformização \sep Sistemas
algorítmicos \sep Educação}
\end{abstract}
\end{portuguese}
\end{polyabstract}

\input{sec-introducao_en.tex}
\input{sec-metodologia_en.tex}
\input{sec-analise_en.tex}
\input{sec-consideracoesfinais_en.tex}

\section*{Funding} This work was carried out with the support of the
Coordination for the Improvement of Higher Education Personnel - Brazil (CAPES)
- Financing Code 001. The study was also carried out with the support of the
project `Percepción y generación de conocimiento sobre los algoritmos de
personalización en las plataformas digitales de comunicación - ALCOM' código
PID2023-148682OA-I00, funded by the Ministry of Science, Innovation and
Universities of Spain.



\printbibliography\label{sec-bib}
%conceptualization,datacuration,formalanalysis,funding,investigation,methodology,projadm,resources,software,supervision,validation,visualization,writing,review
\begin{contributors}[sec-contributors]
\authorcontribution{Willian Fernandes Araujo}[conceptualization,methodology,writing,review]
\authorcontribution{Fernanda Pires de Sá}[methodology,writing]
\end{contributors}

\appendix
\input{anexo_en}

\end{document}
