\documentclass[english]{textolivre}

% metadata
\journalname{Texto Livre}
\thevolume{17}
%\thenumber{1} % old template
\theyear{2024}
\receiveddate{\DTMdisplaydate{2023}{8}{31}{-1}}
\accepteddate{\DTMdisplaydate{2023}{11}{15}{-1}}
\publisheddate{\today}
\corrauthor{Ronaldo Corrêa Gomes Junior}
\articledoi{10.1590/1983-3652.2024.47924}
%\articleid{NNNN} % if the article ID is not the last 5 numbers of its DOI, provide it using \articleid{} commmand 
% list of available sesscions in the journal: articles, dossier, reports, essays, reviews, interviews, editorial
\articlesessionname{dossier}
\runningauthor{Braga and Gomes Junior}
%\editorname{Leonardo Araújo} % old template
\sectioneditorname{Daniervelin Pereira}
\layouteditorname{João Mesquita}

\title{"In the palm of your hand": the ecological and complex nature of teacher agency}
\othertitle{"Na palma da mão": a natureza ecológica e complexa da agência do professor}






\usepackage[inline]{enumitem}
%\usepackage{easyReview}



\author[1]{Junia de Carvalho Fidelis Braga ~\orcid{0000-0002-8450-2061}\thanks{Email: \href{mailto:juniadecarvalhobraga@gmail.com}{juniadecarvalhobraga@gmail.com}}}
\author[1]{Ronaldo Corrêa Gomes Junior~\orcid{0000-0003-4165-8629}\thanks{Email: \href{mailto:ronaldocgomes@gmail.com}{ronaldocgomes@gmail.com}}}
\affil[1]{Universidade Federal de Minas Gerais, Faculdade de Letras, Belo Horizonte, MG, Brasil.}

\addbibresource{article.bib}

\begin{document}
\maketitle
\begin{polyabstract}
\begin{abstract}
The potential of mobile phones as a resource to mediate teaching and learning opportunities has been the subject of research in the educational context. However, the role of these devices in teacher education, especially regarding the exercise of teacher agency, has been insufficiently investigated. The concept of agency has been discussed from different perspectives, and many studies in this area share the idea that agency is one of the characteristics of human behavior. In this paper\footnote{The authors acknowledge the support received from FAPEMIG (Minas Gerais State Agency for Research and Development) in the form of scholarships and financial support.}, we attempt to discuss the concept of agency through the lens of the ecological approach and complexity, perspectives that shed light to the way human beings’ exercise of agency emerges as agents who can influence and be influenced by their context. The aims of this study were to identify: i) instances of agency in the actions of pre-service teachers; ii) actions mediated by the use of mobile phones by these teachers; and iii) intrapersonal issues that can influence the agency of these teachers. Data was collected through a semi-structured questionnaire in undergraduate classes on the use of digital technologies in language teaching and learning. The data was then processed and coded using Atlas-TI software, and the analysis and discussion of the data was based on the characteristics of agency as a complex system. With this study, we hope to demonstrate the multifaceted nature of the concept of agency as these mobile technologies are integrated into teaching practices and to underscore its importance for language-teacher pre-service initial education.



\keywords{Teacher education \sep Mobile technologies \sep Agency \sep Complexity \sep Ecological perspective}
\end{abstract}

\begin{abstract}
O potencial dos celulares como recurso para mediar oportunidades de ensino e aprendizagem tem sido tema de pesquisas no contexto educacional. No entanto, o papel desses dispositivos no contexto de formação, especialmente com relação ao exercício de agência do professor, é tema pouco explorado. O conceito de agência tem sido discutido sob diferentes perspectivas e muitos estudos nessa área compartilham o pensamento de que agência é uma das características do comportamento humano. Neste trabalho, buscamos discutir o conceito de agência pelas lentes da perspectiva ecológica e da complexidade, por se tratar de perspectivas que esclarecem a forma como o exercício da agência pelos seres humanos emerge como agentes que podem influenciar e serem influenciados pelo seu contexto. O objetivo deste estudo foi o de identificar: i) instâncias de agência na atuação de professores em formação inicial; ii) ações mediadas pelo uso de celulares por esses professores; e iii) questões intrapessoais que podem influenciar a agência desses professores. Os dados foram gerados por meio de um questionário semi-estruturado em turmas de graduação e especialização sobre o uso de tecnologias digitais no ensino e aprendizagem de línguas. Esses dados foram tratados e tabulados com o apoio do software Atlas-TI e a análise e discussão dos dados foram pautadas em características da agência como sistema complexo. Com este estudo, esperamos evidenciar a natureza multifacetada do conceito de agência em se tratando da integração de tecnologias móveis na prática docente, destacando a sua importância para a formação inicial de professores de línguas.

\keywords{Formação de professores \sep Tecnologias móveis \sep Agência \sep Complexidadec \sep Perspectiva Ecológica}
\end{abstract}
\end{polyabstract}

\input{sec-introducao.tex}
\input{sec-ecologicalapproach.tex}
\input{sec-mobilelearning.tex}
\input{sec-agency.tex}
\input{sec-thestudy.tex}
\input{sec-dataanalysis.tex}
\input{sec-conclusion.tex}

\printbibliography\label{sec-bib}

%\begin{contributors}[sec-contributors]
%\authorcontribution{Junia de Carvalho Fidelis Braga}[conceptualization,datacuration,formalanalysis,investigation,methodology,projadm,software,validation,visualization,writing,review]
%\authorcontribution{Ronaldo Correa Gomes Junior}[conceptualization,datacuration,formalanalysis,methodology,validation,visualization,writing,review]
%\end{contributors}
\end{document}



