\documentclass[spanish]{textolivre}

% metadata
\journalname{Texto Livre}
\thevolume{17}
%\thenumber{1} % old template
\theyear{2024}
\receiveddate{\DTMdisplaydate{2024}{2}{11}{-1}}
\accepteddate{\DTMdisplaydate{2024}{5}{20}{-1}}
\publisheddate{\today}
\corrauthor{Karla Karina Ruiz Mendoza}
\articledoi{10.1590/1983-3652.2024.51222}
%\articleid{NNNN} % if the article ID is not the last 5 numbers of its DOI, provide it using \articleid{} commmand 
% list of available sesscions in the journal: articles, dossier, reports, essays, reviews, interviews, editorial
\articlesessionname{articles}
\runningauthor{Ruiz Mendoza et al.}
%\editorname{Leonardo Araújo} % old template
\sectioneditorname{Hugo Heredia Ponce}
\layouteditorname{João Mesquita}

\title{Creación y jueceo de ítems: ChatGPT como diseñador y juez}
\othertitle{Criação e julgamento de itens: ChatGPT como designer e juiz}
\othertitle{Item creation and judging: ChatGPT as designer and judge}

\author[1]{Karla Karina Ruiz Mendoza~\orcid{0000-0001-8978-8364}\thanks{Email: \href{mailto:ruiz.karla32@uabc.edu.mx}{ruiz.karla32@uabc.edu.mx}}}
\author[1]{Luis Horacio Pedroza Zúñiga~\orcid{0000-0002-5256-2967}\thanks{Email: \href{mailto:horacio.pedroza@uabc.edu.mx}{horacio.pedroza@uabc.edu.mx}}}
\author[2]{Alma Yadhira López García~\orcid{0000-0002-7474-5799}\thanks{Email: \href{mailto:alma.l@thelearningbar.com}{alma.l@thelearningbar.com}}}
\affil[1]{Universidad Autónoma de Baja, California, IIDE, Ensenada, Baja California, México.}
\affil[2]{TheLearning Bar, Canadá.}

\addbibresource{article.bib}
\usepackage{array,longtable,multirow}
\usepackage{enumitem}


\begin{document}
\maketitle
\begin{polyabstract}
\begin{abstract}
El fin de este estudio fue evaluar la efectividad de la inteligencia
artificial (IA), representada por ChatGPT 4.0, comparada con diseñadores
humanos en la creación de ítems para un examen para el ingreso a la
educación superior en el área de Lengua Escrita. Se utilizó un enfoque
mixto, combinando metodologías clásicas y contemporáneas en evaluación
educativa, incluyendo el juicio de expertos. ChatGPT y cuatro
diseñadores humanos desarrollaron 84 ítems, siguiendo la Taxonomía de
Anderson y Krathwohl para establecer el nivel de demanda cognitiva. Los
ítems fueron evaluados por dos jueces humanos y ChatGPT, utilizando una
rúbrica detallada que incluye claridad, neutralidad, formato, alineación
curricular y redacción. Los resultados mostraron una alta tasa de
aceptación sin cambios tanto para ítems de ChatGPT como para los
humanos, indicando una buena alineación con los estándares de
evaluación. Sin embargo, se observaron diferencias en la necesidad de
cambios menores y mayores propuestos por la rúbrica. El estudio concluye
que tanto la IA como los diseñadores humanos son capaces de generar
ítems de alta calidad, resaltando el potencial de la IA en el diseño de
ítems educativos.

\keywords{Inteligencia Artificial \sep Evaluación educativa \sep
	ChatGPT \sep Diseño de ítems \sep Jueceo}
\end{abstract}

\begin{portuguese}
\begin{abstract}
O objetivo deste estudo foi avaliar a eficácia da inteligência
artificial (IA), representada pelo ChatGPT 4.0, em comparação com
designers humanos na criação de itens para um exame de ingresso ao
ensino superior na área de Língua Escrita. Utilizou-se uma abordagem
mista, combinando metodologias clássicas e contemporâneas em avaliação
educacional, incluindo o julgamento de especialistas. O ChatGPT e quatro
designers humanos desenvolveram 84 itens, seguindo a Taxonomia de
Anderson e Krathwohl para estabelecer o nível de demanda cognitiva. Os
itens foram avaliados por dois juízes humanos e pelo ChatGPT, utilizando
uma rubrica detalhada que inclui clareza, neutralidade, formato,
alinhamento curricular e redação. Os resultados mostraram uma alta taxa
de aceitação sem mudanças tanto para itens do ChatGPT quanto para os
humanos, indicando um bom alinhamento com os padrões de avaliação. No
entanto, foram observadas diferenças na necessidade de mudanças menores
e maiores propostas pela rubrica. Conclui-se que tanto a IA quanto os
designers humanos são capazes de gerar itens de alta qualidade,
destacando o potencial da IA no design de itens educacionais.


\keywords{Inteligência Artificial \sep Avaliação educacional \sep ChatGPT \sep Design de itens \sep Julgamento}
\end{abstract}
\end{portuguese}

\begin{english}
\begin{abstract}
The purpose of this study was to evaluate the effectiveness of
artificial intelligence (AI), represented by ChatGPT 4.0, compared to
human designers in creating items for an exam for entry into higher
education in the area of Written Language. A mixed approach was
utilized, combining classic and contemporary methodologies in
educational evaluation including expert judgment. ChatGPT and four human
designers developed 84 items, following Anderson and
Krathwohl\textquotesingle s Taxonomy to establish the level of cognitive
demand. The items were evaluated by two human judges and ChatGPT, using
a detailed rubric that includes clarity, neutrality, format, curricular
alignment, and writing. The results showed a high rate of acceptance
without changes for both ChatGPT and human items, indicating good
alignment with the evaluation standards. However, differences were
observed in the need for minor and major changes proposed by the rubric.
The study concludes that both AI and human designers are capable of
generating high-quality items, highlighting the potential of AI in the
design of educational items.

\keywords{Artificial Intelligence \sep Educational assessment \sep
	ChatGPT \sep Item design \sep Judging Process}
\end{abstract}
\end{english}

\end{polyabstract}

\input{sec-introduccion}
\input{sec-validezyjueceo}
\input{sec-metodologia}
\input{sec-resultados}
\input{sec-discusion}



\printbibliography\label{sec-bib}
%conceptualization,datacuration,formalanalysis,funding,investigation,methodology,projadm,resources,software,supervision,validation,visualization,writing,review
\begin{contributors}[sec-contributors]
\authorcontribution{Karla Karina Ruiz Mendoza}[investigation,formalanalysis,writing,]
\authorcontribution{Luis Horacio Pedroza Zúñiga}[investigation,supervision,validation,review]
\authorcontribution{Alma Yadhira López García}[methodology,supervision,validation]

\end{contributors}

\appendix

\input{appendix}

\end{document}
