\documentclass[english]{textolivre}

% metadata
\journalname{Texto Livre}
\thevolume{18}
%\thenumber{1} % old template
\theyear{2025}
\receiveddate{\DTMdisplaydate{2024}{7}{31}{-1}}
\accepteddate{\DTMdisplaydate{2024}{11}{29}{-1}}
\publisheddate{\DTMdisplaydate{2025}{4}{17}{-1}}
\corrauthor{Pablo Ramírez Rodríguez}
\articledoi{10.1590/1983-3652.2025.53794}
%\articleid{NNNN} % if the article ID is not the last 5 numbers of its DOI, provide it using \articleid{} commmand 
% list of available sesscions in the journal: articles, dossier, reports, essays, reviews, interviews, editorial
\articlesessionname{dossier}
\runningauthor{Ramírez Rodríguez y Antonov}
%\editorname{Leonardo Araújo} % old template
\sectioneditorname{Hugo Heredia Ponce}
\layouteditorname{João Mesquita}

\title{New challenges and opportunities in technology-assisted phraseology interpreting: the case of Yandex live stream translation}
\othertitle{Novos desafios e oportunidades na interpretação da fraseologia assistida pela tecnologia: o caso da tradução de vídeo ao vivo do Yandex}
%\othertitle{Nuevos retos y oportunidades en la interpretación de la fraseología asistida por la tecnología: el caso de la traducción de vídeo en directo de Yandex}

\author[1]{Pablo Ramírez Rodríguez~\orcid{0000-0002-6168-3736}\thanks{Email: \href{mailto:pablorarod@gmail.com}{pablorarod@gmail.com}}}
\author[2]{Evgeniy Antonov~\orcid{0000-0003-1498-9131}\thanks{Email: \href{mailto:eantonov@kaf65.ru}{eantonov@kaf65.ru}}}
\affil[1]{Universidad de Granada, Granada, Spain.}
\affil[2]{National Research Nuclear University, MEPhI, Moscow, Russia.}

\addbibresource{article.bib}
%\usepackage{easyReview}
% \usepackage[utf8]{inputenc}  % UTF-8 encoding for Unicode characters
% \usepackage[T2A]{fontenc}    % Cyrillic font encoding for Russian characters

\setotherlanguage{russian}   % Define another language
%\usepackage{fontspec}      % For font selection
%\newfontfamily\cyrillicfont[Script=Cyrillic]{Charis SIL}
%\newfontfamily\cyrillicfont{Liberation Serif}
\newfontfamily\cyrillicfontsf{Liberation Sans}

\begin{document}
\maketitle
\begin{polyabstract}
%\begin{spanish}
%\begin{abstract}
%  La interpretación asistida por ordenador (IAO) se ha convertido
%  en un elemento esencial en un mundo cada vez más globalizado e
%  interconectado. Con la creciente demanda de comunicación instantánea en
%  línea y la necesidad de superar las limitaciones lingüísticas en un
%  contexto global, la IAO se ha convertido en una valiosa herramienta para
%  una amplia gama de aplicaciones. Sin embargo, la IAO no está exenta de
%  dificultades. La interpretación de expresiones idiomáticas sigue siendo
%  un obstáculo importante, ya que estas construcciones lingüísticas pueden
%  ser especialmente difíciles de interpretar con precisión debido a su
%  naturaleza culturalmente arraigada. En este contexto, el objetivo de
%  este artículo es abordar el problema de la interpretación de expresiones
%  idiomáticas, en nuestro caso de 52 locuciones verbales en el marco de la
%  IAO, centrándonos en la combinación lingüística español-ruso. El
%  objetivo es analizar cómo esta tecnología responde a los retos de la
%  fraseología idiomática y cómo influye en una comunicación intercultural
%  precisa y eficaz. Para lograr este objetivo, se ha aplicado una
%  metodología exhaustiva que combina el análisis lingüístico con
%  observaciones prácticas de situaciones de interpretación asistida por
%  tecnología en tiempo real utilizando el traductor de voz en directo y
%  los contenidos multimedia proporcionados por Yandex. Los resultados de
%  esta investigación aportan una comprensión más profunda de cómo la
%  tecnología de la traducción y la interpretación aborda los retos de la
%  expresión idiomática, al tiempo que proporcionan una visión crítica de
%  la eficacia y las limitaciones de las soluciones tecnológicas en el
%  ámbito de la comunicación intercultural.
%  
%
%
%\keywords{IAO \sep Fraseología \sep Locuciones verbales \sep Traducción en
%directo \sep Interpretación automática}
%\end{abstract}
%\end{spanish}

\begin{english}
\begin{abstract}
Computer-assisted interpreting (CAI) has become an essential
element in an increasingly globalized and interconnected world. With the
growing demand for instant online communication and the need to overcome
language limitations in a global context, CAI has become a valuable tool
for a wide range of applications. However, the CAI is not without its
challenges. The interpretation of idiomatic expressions remains a
significant barrier, as these linguistic constructs can be particularly
difficult to interpret accurately due to their culturally embedded
nature. In this context, the objective of this article is to address the
problem of interpreting idiomatic expressions, in our case 52 verbal
idioms within the framework of CAI, focusing on the Spanish-Russian
language combination. The aim is to analyze how this technology meets
the challenges of idiomatic phraseology and how it influences accurate
and effective intercultural communication. To achieve this goal, a
comprehensive methodology has been applied, combining linguistic
analysis with practical observations of real-time technology-assisted
interpreting situations using the live on-air speech translator and
multimedia content provided by Yandex. The results of this research
provide a deeper understanding of how translation and interpreting
technology addresses the challenges of idiomatic expression, while also
providing critical insight into the effectiveness and limitations of
technological solutions in the field of intercultural communication.

\keywords{CAI \sep Phraseology \sep Verbal idioms \sep Live stream translation \sep
Automatic interpreting} 
\end{abstract}
\end{english}

\begin{portuguese}
\begin{abstract}
  A interpretação assistida por computador (IAC) tornou-se um
  elemento essencial em um mundo cada vez mais globalizado e
  interconectado. Com a crescente demanda por comunicação instantânea
  online e a necessidade de superar as limitações linguísticas em um
  contexto global, a IAC se tornou uma ferramenta valiosa para uma ampla
  gama de aplicações. No entanto, a IAC não está isenta de dificuldades. A
  interpretação de expressões idiomáticas continua sendo um obstáculo
  significativo, pois essas construções linguísticas podem ser
  especialmente difíceis de interpretar com precisão devido à sua natureza
  culturalmente enraizada. Nesse contexto, o objetivo deste artigo é
  abordar o problema da interpretação de expressões idiomáticas, no nosso
  caso de 52 locuções verbais no âmbito da IAC, focando na combinação
  linguística espanhol-russo. O objetivo é analisar como essa tecnologia
  responde aos desafios da fraseologia idiomática e como influencia uma
  comunicação intercultural precisa e eficaz. Para alcançar esse objetivo,
  foi aplicada uma metodologia exaustiva que combina a análise linguística
  com observações práticas de situações de interpretação assistida por
  tecnologia em tempo real, utilizando o tradutor de voz ao vivo e os
  conteúdos multimídia fornecidos pela Yandex. Os resultados desta
  pesquisa proporcionam uma compreensão mais profunda de como a tecnologia
  de tradução e interpretação enfrenta os desafios da expressão
  idiomática, ao mesmo tempo que oferecem uma visão crítica da eficácia e
  das limitações das soluções tecnológicas no âmbito da comunicação
  intercultural.
  
\keywords{IAC \sep Fraseologia \sep Locuções verbais \sep Tradução em tempo
real \sep Interpretação automática}
\end{abstract}
\end{portuguese}


\end{polyabstract}

\input{sec-intro.tex}
\input{sec-theoretical.tex}
\input{sec-methodology.tex}
\input{sec-results.tex}
\input{sec-conclusion.tex}

\printbibliography\label{sec-bib}
%conceptualization,datacuration,formalanalysis,funding,investigation,methodology,projadm,resources,software,supervision,validation,visualization,writing,review
\begin{contributors}[sec-contributors]
\authorcontribution{Pablo Ramírez Rodríguez}[conceptualization,datacuration,writing,investigation,resources,review,supervision,validation]
\authorcontribution{Evgeniy Antonov}[formalanalysis,datacuration,writing,methodology,resources,supervision,visualization]
\end{contributors}
\end{document}
