\documentclass[english]{textolivre}

% metadata
\journalname{Texto Livre}
\thevolume{17}
%\thenumber{1} % old template
\theyear{2024}
\receiveddate{\DTMdisplaydate{2024}{1}{2}{-1}}
\accepteddate{\DTMdisplaydate{2024}{2}{18}{-1}}
\publisheddate{\DTMdisplaydate{2024}{5}{20}{-1}}
\corrauthor{Francisco Javier Olivar de Julián}
\articledoi{10.1590/1983-3652.2024.49369}
%\articleid{NNNN} % if the article ID is not the last 5 numbers of its DOI, provide it using \articleid commmand 
% list of available sesscions in the journal: articles, dossier, reports, essays, reviews, interviews, editorial
\articlesessionname{articles}
\runningauthor{Olivar de Julián}
%\editorname{Leonardo Araújo} % old template
\sectioneditorname{Daniervelin Pereira}
\layouteditorname{João Mesquita}

\title{Automated journalism in news about traffic accidents: emergency services information dump}
\othertitle{Jornalismo automatizado em notícias sobre acidentes de trânsito: despejo de informações de serviços de emergência}

\author[1]{Francisco Javier Olivar de Julián~\orcid{0000-0002-2030-2458}\thanks{Email: \href{mailto:franciscojavier.olivar@unir.net}{franciscojavier.olivar@unir.net}}}

\affil[1]{Universidad Internacional de La Rioja, Facultad de Empresa y Comunicación / ESIT, Logroño, La Rioja, España.}

\addbibresource{article.bib}

\begin{document}
\maketitle
\begin{polyabstract}
\begin{abstract}
In this work, the process of producing news about
events in the Spanish digital press has been studied. A content analysis
was carried out on news related to the main external causes of mortality
(traffic accidents, accidental falls, drowning and suicides) published
in the six main Spanish digital media (\url{elpais.com}, \url{ elmundo.es}, \url{abc.es}, \url{lavanguardia.com}, \url{elconfidencial.com} and \url{20minutos.es})
during the period 2010-2020. The selected news items ($n=5,727$) were
obtained through the digital newspaper library \emph{Mynewsonline}. A
qualitative study has also been carried out (in-depth interviews with
professionals and a focus group). The results show a high number of news
items published on traffic compared to other external causes of death
with higher mortality. It is also appreciated that the media
 automatically transfer the information they receive from the emergency
services. It has even been detected that most of the news events coded
in this study are exact copies that are published on the same day in
different media (headlines and the body of the news).

\keywords{Automated journalism \sep News \sep Traffic accidents \sep Digital press \sep Mynewsonline}
\end{abstract}

\begin{portuguese}
\begin{abstract}
Neste trabalho foi estudado o processo de produção de
notícias sobre acontecimentos na imprensa digital espanhola. Foi
realizada uma análise de conteúdo de notícias relacionadas às principais
causas externas de mortalidade (acidentes de trânsito, quedas
acidentais, afogamentos e suicídios) publicadas nos seis principais
meios de comunicação digitais espanhóis (\url{elpais.com}, \url{ elmundo.es}, \url{abc.es}, \url{lavanguardia.com}, \url{elconfidencial.com} e \url{20minutos.es})
durante o período 2010-2020. As notícias selecionadas ($n=5.727$) foram
obtidas por meio da joalheria digital \emph{Mynewsonline}. Também foi
realizado um estudo qualitativo (entrevistas em profundidade com
profissionais e grupo focal). Os resultados mostram elevado número de
notícias publicadas sobre trânsito em comparação com outras causas
externas de morte com maior mortalidade. Também é apreciado que os meios
de comunicação transfiram automaticamente as informações que recebem dos
serviços de emergência. Foi ainda detectado que a maior parte das
notícias codificadas neste estudo são cópias exatas que são publicadas
no mesmo dia em diferentes meios de comunicação (manchetes e corpo da
notícia).

\keywords{Jornalismo automatizado \sep Notícias \sep Acidentes de trânsito \sep Imprensa digital \sep Mynewsonline}
\end{abstract}
\end{portuguese}
\end{polyabstract}

\input{sec-introduction.tex}
\input{sec-theoreticalbackground.tex}
\input{sec-methodology.tex}
\input{sec-results.tex}
\input{sec-conclusions.tex}


\printbibliography\label{sec-bib}

\end{document}
