% !TEX program = xelatex
% !TEX encoding = UTF-8 Unicode
% !BIB program = biber
\documentclass[english]{textolivre}



% metadata
\journalname{Texto Livre}
\thevolume{17}
%\thenumber{1} % old template
\theyear{2024}
\receiveddate{\DTMdisplaydate{2024}{5}{21}{-1}}
\accepteddate{\DTMdisplaydate{2024}{6}{22}{-1}}
\publisheddate{\today}
\corrauthor{Hussein Abu-Rayyash}
\articledoi{10.1590/1983-3652.2024.52687}
%\articleid{NNNN} % if the article ID is not the last 5 numbers of its DOI, provide it using \articleid{} commmand 
% list of available sesscions in the journal: articles, dossier, reports, essays, reviews, interviews, editorial
\articlesessionname{articles}
\runningauthor{Abu-Rayyash; Alhawamdeh and Ringomon}
%\editorname{Leonardo Araújo} % old template
\sectioneditorname{Daniervelin Pereira}
\layouteditorname{João Mesquita}

\title{The eye-ear relationship: investigating auditory impacts on subtitle reading and comprehension}
\othertitle{A relação olho-ouvido: investigando os impactos auditivos na leitura e compreensão de legendas}

\author[1]{Hussein Abu-Rayyash ~\orcid{0000-0002-9695-4030  }\thanks{Email: \href{mailto:haburayy@kent.edu }{haburayy@kent.edu }}}
\author[2]{Shatha Alhawamdeh~\orcid{0009-0001-6342-7242}\thanks{Email: \href{mailto:salhawam@kent.edu}{salhawam@kent.edu}}}
\author[1]{Yuri Ringomon~\orcid{0009-0000-6508-8769}\thanks{Email: \href{mailto:yringomo@kent.edu }{yringomo@kent.edu }}}
\affil[1]{Kent State University, College of Arts and Sciences, Department of Modern \& Classical Language Studies, Kent, Ohio, USA.}
\affil[2]{College College of Arts and Sciences Department of Middle East and North Africa Studies Oberlin, Ohio, USA.}

\addbibresource{article.bib}
\usepackage{tabularx,multirow}%,makecell}%,multirow}

\newcolumntype{Y}{>{\centering\arraybackslash}X}
%\usepackage{longtable}

\setotherlanguage{arabic}
\newfontfamily\arabicfont[Script=Arabic]{Amiri}
\newfontfamily\arabicfontsf[Script=Arabic]{Amiri}
\newfontfamily\arabicfonttt[Script=Arabic]{Amiri}

\begin{document}
\maketitle
\begin{polyabstract}
\begin{abstract}
  The study employs eye-tracking measures and comprehension
  quizzes to investigate linguistic processing during subtitle reading and
  comprehension in Arabic and Spanish. Specifically, it examines the
  impact of native English (L1) audio, foreign Spanish or Arabic (L2)
  audio, or the absence of audio on subtitle reading strategies and the
  processing of embedded linguistic units in novice learners. The study
  utilizes a 2x3 within-subjects design, manipulating audio condition (L1
  audio, L2 audio, no audio) and subtitle language (Spanish, Arabic).
  Global eye-tracking analyses reveal subtitle reading patterns, while
  local analyses focus on the processing of specific linguistic units.
  Comprehension quizzes assess learning outcomes after the viewing of the
  subtitled videos. The findings suggest that the absence of audio leads
  to increased fixation durations, indicating greater attentional demand.
  Spanish subtitles result in higher fixation counts and prolonged reading
  times compared to Arabic, reflecting deeper cognitive processing. The
  study highlights the complex interplay between audio presence, subtitle
  language, and individual proficiency in shaping attentional allocation
  and lexical processing during language acquisition through audio-visual
  media.
\keywords{Eye Movement \sep Audio \sep Subtitles \sep Reading \sep Comprehension}
\end{abstract}

\begin{abstract}
O estudo emprega medidas de rastreamento ocular e
questionários de compreensão para investigar o processamento linguístico
durante a leitura e compreensão de legendas em árabe e espanhol.
Especificamente, examina o impacto do áudio nativo em inglês (L1), áudio
estrangeiro em espanhol ou árabe (L2) ou a ausência de áudio nas
estratégias de leitura de legendas e no processamento de unidades
linguísticas inseridas em aprendizes iniciantes. O estudo utiliza um
delineamento intra-sujeitos 2x3, manipulando a condição de áudio (áudio
L1, áudio L2, sem áudio) e a língua das legendas (espanhol, árabe). As
análises globais de rastreamento ocular revelam padrões de leitura de
legendas, enquanto as análises locais se concentram no processamento de
unidades linguísticas específicas. Os questionários de compreensão
avaliam os resultados de aprendizagem após assistir aos vídeos
legendados. Os resultados sugerem que a ausência de áudio leva a um
aumento na duração das fixações, indicando uma maior demanda atencional.
As legendas em espanhol resultam em um maior número de fixações e tempos
de leitura prolongados em comparação com o árabe, refletindo um
processamento cognitivo mais profundo. O estudo destaca a complexa
interação entre a presença de áudio, a língua das legendas e a
proficiência individual na alocação de atenção e no processamento
lexical durante a aquisição de línguas através de mídia audiovisual.

\keywords{Movimento Ocular \sep Áudio \sep Legendas \sep Leitura \sep Compreensão}
\end{abstract}


\end{polyabstract}

\input{introduction.tex}
\input{literaturereview.tex}
\input{methodology.tex}
\input{results.tex}
\input{discussion.tex}
\input{conclusion.tex}

\printbibliography\label{sec-bib}
%conceptualization,datacuration,formalanalysis,funding,investigation,methodology,projadm,resources,software,supervision,validation,visualization,writing,review

\section*{Funding Statement}

This research received grant no. (170/2023) from the Arab Observatory for Translation (an affiliate of ALECSO), which is supported by the Literature, Publishing \& Translation Commission in Saudi Arabia.

\section*{Acknowledgementle}

We sincerely thank Dr. Sixin Liao for her invaluable feedback, which significantly contributed to refining the ideas, strengthening the analysis, and improving the overall quality of this research.  

\begin{contributors}[sec-contributors]
	
\authorcontribution{Hussein Abu-Rayyash }[conceptualization,methodology,software,investigation,supervision,writing]
\authorcontribution{Shatha Alhawamdeh}[conceptualization,writing,validation,investigation]
\authorcontribution{Yuri Ringomon}[conceptualization,validation,writing]
\end{contributors}
\end{document}
