\documentclass[portuguese]{textolivre}

% metadata
\journalname{Texto Livre}
\thevolume{18}
%\thenumber{1} % old template
\theyear{2025}
\receiveddate{\DTMdisplaydate{2024}{10}{25}{-1}}
\accepteddate{\DTMdisplaydate{2025}{2}{4}{-1}}
\publisheddate{\today}
\corrauthor{Gabriel Nicolau Souza}
\articledoi{10.1590/1983-3652.2025.55494}
%\articleid{NNNN} % if the article ID is not the last 5 numbers of its DOI, provide it using \articleid{} commmand 
% list of available sesscions in the journal: articles, dossier, reports, essays, reviews, interviews, editorial
\articlesessionname{articles}
\runningauthor{Souza e Silva}
%\editorname{Leonardo Araújo} % old template
\sectioneditorname{Daniervelin Pereira}
\layouteditorname{João Mesquita}

\title{Videoaulas e produção de artigo acadêmico: o que é ensinado?}
\othertitle{Videolessons and production of the academic article: what is taught?}

\author[1]{Gabriel Nicolau de Souza~\orcid{0009-0007-2591-6321}\thanks{Email: \href{mailto:gabrielnicolauestudante@gmail.com}{gabrielnicolauestudante@gmail.com}}}
\author[1]{Elizabeth Maria da Silva~\orcid{0000-0002-1355-493X}\thanks{Email: \href{mailto:elizabeth.maria@professor.ufcg.edu.br}{elizabeth.maria@professor.ufcg.edu.br}}}

\affil[1]{Universidade Federal de Campina Grande, Academia de Letras, Campina Grande, PB, Brasil.}


\addbibresource{article.bib}

%\usepackage{array,amssymb,calc,longtable,ulem}
\usepackage{longtable}

\usepackage{nameref}

%\makeatletter
%\let\orgdescriptionlabel\descriptionlabel
%\renewcommand*{\descriptionlabel}[1]{%
%  \let\orglabel\label
%  \let\label\@gobble
%  \phantomsection
%  \edef\@currentlabel{#1}%
%  %\edef\@currentlabelname{#1}%
%  \let\label\orglabel
%  \orgdescriptionlabel{#1}%
%}
%\makeatother

\makeatletter
\newcounter{descitem}
\let\orgdescriptionlabel\descriptionlabel
\renewcommand*{\descriptionlabel}[1]{%
  \let\orglabel\label
  \let\label\@gobble
  \phantomsection
  \refstepcounter{descitem}
  \edef\@currentlabel{\thedescitem}%
  \let\label\orglabel
  \orgdescriptionlabel{#1}%
}
\makeatother


\begin{document}
\maketitle
\begin{polyabstract}
\begin{abstract}
Neste trabalho, objetiva-se identificar e analisar
objetos de ensino explorados em videoaulas sobre produção de artigo
acadêmico. Metodologicamente, adota-se o método quanti-qualitativo, a
partir da netnografia. Como aporte teórico, apresentam-se práticas de
ensino de Língua Portuguesa em contexto didático-digital \cite{laurentino2023}, artigo acadêmico \cite{motta-roth2010}, as etapas de
produção textual \cite{antunes2003} e objetos de ensino \cite{linodearaujo2014}. Os dados constam de quinze videoaulas publicadas na plataforma
YouTube e destacam o foco na estrutura composicional do gênero como o
objeto de ensino mais recorrente. Observa-se que essa recorrência
decorre das especificidades do contexto didático-digital que abriga as
videoaulas, justificando seu modelo de produção e método de ensino.
Nesta análise, dá-se, assim, visibilidade a novos meios de ensino de
gêneros acadêmicos, como o artigo, e às formas pelas quais os objetos de
ensino são explorados em videoaulas.
 
\keywords{Contexto didático-digital \sep Videoaulas \sep Artigo
acadêmico \sep Objetos de ensino \sep Etapas de produção textual}
\end{abstract}

\begin{english}
\begin{abstract}
  This study aims to identify and analyze teaching
  components explored in videolessons about academic article production.
  Methodologically, the quantitative-qualitative method is adopted, based
  on netnography. As a theoretical contribution, Portuguese language
  teaching practices in a didactic-digital context are presented
  \cite{laurentino2023}, as well as academic article \cite{motta-roth2010}, stages of textual production \cite{antunes2003} and
  teaching components \cite{linodearaujo2014}. The data are contained in
  fifteen videolessons on the YouTube platform and highlight the focus on
  the compositional structure of the genre as the most recurrent teaching
  component. It is observed that this recurrence stems from the
  specificities of the didactic-digital context that houses the
  videolessons, justifying their production model and teaching method. In
  this analysis, visibility is thus given to new means of teaching
  academic genres, such as the article, and to the ways in which teaching
  components are explored in videolessons.
  
\keywords{Didactic-digital context \sep Videolessons \sep Academic
article \sep Teaching components \sep Stages of textual production}
\end{abstract}
\end{english}
\end{polyabstract}

\input{sec-intro.tex}
\input{sec-embasamento.tex}
\input{sec-metodologia.tex}
\input{sec-objetos.tex}
\input{sec-consideracoes.tex}

\printbibliography\label{sec-bib}
%conceptualization,datacuration,formalanalysis,funding,investigation,methodology,projadm,resources,software,supervision,validation,visualization,writing,review
\begin{contributors}[sec-contributors]
\authorcontribution{Gabriel Souza}[datacuration,formalanalysis,investigation,methodology,projadm,resources,visualization,writing,review]
\authorcontribution{Elizabeth Silva}[ceonptualization,formalanalysis,methodology,projadm,resources,supervision,validation,visualization,writing,review]
\end{contributors}
\end{document}
