\documentclass[english]{textolivre}

% metadata
\journalname{Texto Livre}
\thevolume{17}
%\thenumber{1} % old template
\theyear{2024}
\receiveddate{\DTMdisplaydate{2024}{3}{20}{-1}}
\accepteddate{\DTMdisplaydate{2024}{4}{23}{-1}}
\publisheddate{\today}
\corrauthor{María Bobadilla-Pérez}
\articledoi{10.1590/1983-3652.2024.51692}
%\articleid{NNNN} % if the article ID is not the last 5 numbers of its DOI, provide it using \articleid{} commmand 
% list of available sesscions in the journal: articles, dossier, reports, essays, reviews, interviews, editorial
\articlesessionname{articles}
\runningauthor{Galán-Rodríguez; Bobadilla-Pérez; Barros-Grela}
%\editorname{Leonardo Araújo} % old template
\sectioneditorname{Hugo Heredia Ponce}
\layouteditorname{João Mesquita}

\title{Attitudes and perceptions: the role of artificial intelligence in the training of future secondary school foreign language teachers}
\othertitle{Atitudes e percepções: o papel da inteligência artificial na formação de futuros professores de línguas estrangeiras do ensino secundário}
\author[1]{Noelia Mª Galán-Rodríguez~\orcid{0000-0001-6662-7269}\thanks{Email: \href{mailto:noelia.galan@udc.es}{noelia.galan@udc.es}}}
\author[1]{María Bobadilla-Pérez~\orcid{0000-0002-4972-5980}\thanks{Email: \href{mailto:m.bobadilla@udc.es}{m.bobadilla@udc.es}}}
\author[2]{Eduardo Barros-Grela~\orcid{0000-0002-7533-5580}\thanks{Email: \href{mailto:ebarros@udc.es}{ebarros@udc.es}}}
\affil[1]{Universidade da Coruña, College of Education, Department of Specific Teaching Training and Research and Diagnosis Methods in Education, A Coruña, Spain.}
\affil[2]{Universidade da Coruña, College of Arts and Letters, Department of English, A Coruña, Spain.}

\addbibresource{article.bib}

\newcommand\subsubsubsection[1]{%
  \paragraph{#1}\mbox{}\\
}

\begin{document}
\maketitle
\begin{polyabstract}
\begin{abstract}
This study explores prospective foreign language teachers' attitudes towards integrating artificial intelligence (AI) into their teaching practices, focusing on the TWEE tool. Using a mixed-methods approach, data from 29 pre-service secondary school teachers in Spain were analyzed using SWOT analysis to assess TWEE's strengths, weaknesses, opportunities, and threats. Internal strengths include resource variety and practicality, while weaknesses involve interface issues and language constraints. External opportunities include quick activity creation, while threats center on ethical concerns and teacher redundancy risks. Despite concerns, participants generally express a positive attitude towards AI, acknowledging its potential to enhance written tasks, motivation, and technological literacy. Pedagogical implications underscore the need for a balanced AI integration approach, considering accessibility and ethical concerns. Further research is suggested to explore AI tool implementation in classrooms. This study contributes to understanding AI's role in language education and informs strategic planning for its integration.


\keywords{Foreign language teaching \sep Artificial intelligence \sep Secondary Education \sep TWEE \sep Pedagogical perspectives}
\end{abstract}

\begin{portuguese}
\begin{abstract}
Este estudo explora as atitudes dos futuros professores de línguas estrangeiras relativamente à integração da inteligência artificial (IA) nas suas práticas de ensino, centrando-se na ferramenta TWEE. Utilizando uma abordagem de métodos mistos, os dados de 29 professores de ensino secundário em início de carreira em Espanha foram analisados utilizando a análise SWOT para avaliar os pontos fortes, os pontos fracos, as oportunidades e as ameaças da TWEE. Os pontos fortes internos incluem a variedade de recursos e o carácter prático, enquanto os pontos fracos envolvem questões de interface e restrições linguísticas. As oportunidades externas incluem a criação rápida de atividades, enquanto as ameaças se centram em preocupações éticas e riscos de redundância de professores. Apesar das preocupações, os participantes expressam geralmente uma atitude positiva em relação à IA, reconhecendo o seu potencial para melhorar as tarefas escritas, a motivação e o letramento tecnológico. As implicações pedagógicas sublinham a necessidade de uma abordagem equilibrada da integração da IA, tendo em conta a acessibilidade e as preocupações éticas. Sugere-se mais investigação para explorar a implementação de ferramentas de IA nas salas de aula. Este estudo contribui para a compreensão do papel da IA no ensino das línguas e serve de base ao planeamento estratégico para a sua integração.

\keywords{Ensino de línguas estrangeiras \sep Inteligência artificial \sep Ensino secundário \sep TWEE \sep Perspectivas pedagógicas}
\end{abstract}
\end{portuguese}
\end{polyabstract}

\input{sec-introduction}
\input{sec-theoreticalframework}
\input{sec-artificialintelligencein}
\input{sec-methodology}
\input{sec-discussion}
\input{sec-conclusion}

\printbibliography\label{sec-bib}

\begin{contributors}[sec-contributors]
\authorcontribution{Noelia María Galán-Rodríguez}[datacuration,formalanalysis,investigation]
\authorcontribution{María Bobadilla-Pérez}[conceptualization,methodology,writing,validation]
\authorcontribution{Eduardo Barros-Grela}[review,projadm,supervision,visualization]
\end{contributors}

\end{document}
