\documentclass[english]{textolivre}

% metadata
\journalname{Texto Livre}
\thevolume{17}
%\thenumber{1} % old template
\theyear{2024}
\receiveddate{\DTMdisplaydate{2024}{3}{18}{-1}}
\accepteddate{\DTMdisplaydate{2024}{6}{25}{-1}}
\publisheddate{\today}
\corrauthor{Liudmila Shafirova}
\articledoi{10.1590/1983-3652.2024.51663}
%\articleid{NNNN} % if the article ID is not the last 5 numbers of its DOI, provide it using \articleid{} commmand 
% list of available sesscions in the journal: articles, dossier, reports, essays, reviews, interviews, editorial
\articlesessionname{articles}
\runningauthor{Shafirova and Araújo e Sá}
%\editorname{Leonardo Araújo} % old template
\sectioneditorname{Daniervelin Pereira}
\layouteditorname{João Mesquita}

\title{Students making videos on social media: exploring the potential of online videos for language learning}
\othertitle{Produção de vídeos em redes sociais por estudantes: explorando o potencial de vídeos online para a aprendizagem de línguas}

\author[1]{Liudmila Shafirova~\orcid{0000-0003-3743-2029}\thanks{Email: \href{mailto:liudmila.shafirova@ua.pt }{liudmila.shafirova@ua.pt }}}
\author[1]{Maria Helena Araújo e Sá~\orcid{0000-0002-6623-9642}\thanks{Email: \href{mailto:helenasa@ua.pt}{helenasa@ua.pt}}}
	
\affil[1]{University of Aveiro, Department of Education and Psychology, CIDTFF, Aveiro, Portugal.}

\addbibresource{article.bib}

\begin{document}
	
\maketitle
\begin{polyabstract}

\begin{abstract}
In recent years, communication on social media has undergone major changes, shifting towards being more multimodal and video-centred. This study investigates the engagement of Portuguese university students in video production within social media platforms and examines their perceptions of language learning opportunities inherent to this practice. We conducted a questionnaire among 212 students at the University of Aveiro (Portugal), delving into the role of video-making as an interactive learning tool. Our results show that roughly 22\% of the respondents make videos, indicating that video production is not very common among them, especially when compared to video watching. Despite this, our analysis highlights several possible educational benefits of making online videos. Our data illustrate the interactive and multilingual nature of online video-making, encompassing practices such as searching for information, collaborating with others, or watching other similar videos in different languages. Finally, while this study highlights the infrequency of video production, it underscores its potential as a holistic approach to language learning and the development of multiliteracies.


\keywords{Online videos \sep Social media \sep Multiliteracies \sep Informal language learning}
\end{abstract}

\begin{portuguese}
\begin{abstract}
Nos últimos anos, a comunicação nas redes sociais passou por significativas transformações, evoluindo para uma abordagem mais multimodal e orientada para o vídeo. Este estudo investiga o envolvimento de estudantes universitários portugueses na produção de vídeos em plataformas de redes sociais e examina as suas percepções de oportunidades de aprendizagem de línguas inerentes a esta atividade social. Para atingir estes objetivos, realizamos um questionário com 212 estudantes da Universidade de Aveiro. Os resultados mostram que cerca de 22\% dos inquiridos fazem vídeos, o que indica que esta atividade não é muito comum entre eles, especialmente quando comparada com a visualização de vídeos. Apesar disso, os resultados destacam vários benefícios educativos possíveis da criação de vídeos online, como a sua natureza interativa e multilíngue. Essas práticas incluem a busca de informações, a colaboração com outros e a visualização de vídeos semelhantes em diferentes línguas. Por fim, embora este estudo destaque a pouca frequência da produção de vídeos, sublinha o seu potencial como abordagem holística à aprendizagem de línguas e ao desenvolvimento de multiletramentos.

\keywords{vídeos online \sep redes sociais \sep multiletramentos \sep aprendizagem de línguas fora da escola}
\end{abstract}
\end{portuguese}
\end{polyabstract}

\input{sec-introduction.tex}%
\input{sec-theoframework.tex}%
\input{sec-materialsand.tex}%
\input{sec-results.tex}%
\input{sec-discussion.tex}

\section*{Funding}\label{sec-funding}
This work is funded by national funds through FCT – Fundação para a Ciência e a Tecnologia,
I.P., under the Scientific Employment Stimulus - Individual Call – [2022.06443.CEECIND] with DOI: 
10.54499/\allowbreak{}2022.06443.\allowbreak{}CEECIND/\allowbreak{}CP1720/\allowbreak{}CT0040 
and the CIDTFF Research Centre (projects UIDB/00194/2020 and UIDP/00194/2020).
It was also partly supported by OralGrab. Grabar vídeos y audios para ensenyar y aprender (PID2022-141511NB-100, Ministry of Science and Innovation), Gr@el - Grup de recerca sobre aprenentatge i ensenyament de llengües (SGR471, AGAUR), and by DEFINERS: Digital language learning of language teachers (TED2021-129984A-I00, Ministry of Science and Innovation, Spain).


\printbibliography
\label{sec-bib}
%conceptualization,datacuration,formalanalysis,funding,investigation,methodology,projadm,resources,software,supervision,validation,visualization,writing,review
\begin{contributors}[sec-contributors]
\authorcontribution{Liudmila Shafirova}[conceptualization,datacuration,formalanalysis,investigation,methodology,projadm,resources,visualization,writing,review]
\authorcontribution{Helena Araújo e Sá}[conceptualization,funding,funding,investigation,methodology,supervision,validation,writing,review]
\end{contributors}


\newpage
\input{appendix}

\end{document}
